---
title: 'Reproducible Research: Peer Assessment 1'
output: html_document
---

### Get raw data
Load the data and Process/transform the data into a format suitable for analysis.

#### Set WD
```{r}
setwd("C://temp/")
```

#### Load required libraries
```{r, message=F, warning=F }
library(dplyr)
library(ggplot2)
library(lattice)
```

#### Load data base from working directory
```{r}
activity <- data.frame(read.csv("activity.csv"),stringsAsFactors = FALSE)
```

## 1-Calculate the total number of steps taken per day

```{r,echo=FALSE }
activity1 <- activity %>% 
    na.omit() %>% 
    summarise(Total_Steps = sum(steps))

activity1
```

## 2-Histogram of the total number of steps taken each day

```{r,echo=FALSE ,message=F, warning=F}
activity2 <- activity %>% 
    na.omit() %>% 
    select(date,steps) %>%
    group_by(date) %>%
    summarise(step_sum = sum(steps))
   

ggplot(activity2, aes(x=step_sum)) +
    geom_histogram(colour="black", fill="white")
```

##3-Summary Statistics of the total number of steps taken per day

```{r,echo=FALSE}
summary(activity2)
```

##4-Time series plot of 5-minute intervals and the average number of steps taken, averaged across all days.

```{r,echo=FALSE ,message=F, warning=F}
activity4 <- activity %>% 
    na.omit() %>% 
    select(steps, interval) %>%
    group_by(interval) %>%
    summarise(ts = mean(steps))

plot(activity4$interval, activity4$ts, type = "l", xlab = "5-min interval", 
     ylab = "Average across all Days", main = "Time Series- Average Steps", 
     col = "blue")
```

## 5-Total number of missing values in the dataset.

```{r,echo=FALSE }
activity5 <- activity %>% 
    filter(is.na(steps))%>%
    summarize(NA_Number = n())

activity5
```

### 6-New dataset with missing data filled in.
```{r, echo=FALSE }
activity6 <- activity %>%
     mutate(steps= ifelse(is.na(steps), mean(steps, na.rm=TRUE), steps))
```

## 7-Histogram and summary statistics of the total number of steps taken each day with the none missing data.

```{r,echo=FALSE ,message=F, warning=F}
activity7 <- activity6 %>% 
    select(date,steps) %>%
    group_by(date) %>%
    summarise(step_sum = sum(steps))


ggplot(activity7, aes(x=step_sum)) +
    geom_histogram(colour="black", fill="white")

summary(activity7)
```

#### Once missing values have been replaced with the total step mean,there is little difference in the summary statistics for both data sets.

### 8-Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r,echo=FALSE ,message=F, warning=F}
activity8 <- activity6 %>%
    mutate(weekday= weekdays(as.Date(date)))%>%
    mutate(wday = ifelse((weekday == 'Saturday'| weekday == 'Sunday'),
            "Weekend","Weekday"))%>%
    group_by(wday,interval) %>%
    summarise(ts = mean(steps))
```

## 9-Time series plot of 5-minute intervals and the average number of steps taken, averaged across all weekday days or weekend days. 


```{r,echo=FALSE ,message=F, warning=F}
xyplot(ts ~ interval | wday, activity8, type = "l", layout = c(1, 2), 
       xlab = "Interval", ylab = "Number of steps")
```

#### There are differences between the two step disributions when comparing weekend and weekday.



